Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-08-28 chatbot_classification/aaai_2021/paper.tex
modified:   paper.pdf   modified:   paper.tex

  Task-specific  conversational  chatbot  \cite{wen2016network}  are  designed  to
  share   the  working  pressure  of  human  customer  service  agents  who  are
  responsible for solving customers' questions or queries about certain products
  or  services.  Regardless  of  whether  the  conversation  is  single-turn  or
  multi-turn,  the  essential  technical  solution  behind the chatbot is intent
  classification model which can identify the correct intent behind user's input
  text and find the corresponding answer.

  Although  task-specific  conversational chatbot already has a wide application
  in various business and industries, it's still a quite challenging task due to
  its  natural  properties of low-resource. First, customers' utterance within a
  conversation  is  usually  quite  short  and  completed  in a few words. Short
  text \cite{song2014short}  are generally more ambiguous in comparison with long
  texts  such  as  paragraphs  or  documents  since  they  don't  contain enough
  contextual  information,  which poses a great challenge \cite{chen2019deep} for
  classification        task \cite{phan2008learning,yan2009dynamic,hua2015short}.
  Second,  at  the  initial  stage  of  building  a chatbot for specific task or
  service, it's often extremely hard to collect sufficient data samples for each
  class.  The  reason  is  that it's usually too expensive and time consuming to
  extract  different  ways  of natural language expression for each intent class
  from  history  conversation  log  or  even  manually  compose  the corpus from
  nothing.  Therefore,  the  key  to  building a task-specific chatbot with high
  performance     becomes     solving     the     challenge     of    short-text
  classification \cite{sriram2010short}        problem       under       few-shot
  setting \cite{yu2018diverse}.

@@

  Despite the success of these 3 approaches, they still have some limitations in
  task-specific chatbot scenario. As for the text classification model approach,
  it's   quite   hard  to  use  data  augmentation  method  to  solve  the  data
  insufficiency  challenge since the it's usually unfeasible to get large amount
  of  domain-specific  data  when  facing  a  new  task. As for label-word joint
  embedding  approach,  it  poses  noticeable requirements on label information.
  However,  in  our  cases,  since  we have large amount of class labels without
  being well-defined for question answering chatbot, the only thing we can do is
  to  choose  one  of the user utterance as the class label to represent the key
  information  of  the  whole class. In this way, the label will become not only
  prolix  but  also  biased,  which  will  bring  some  negative effect to label
  embedding.  With  respect  to  the sentence-pair model approach, though we can
  obtain   large   amount  of  data  in  semantically  duplicate  sentence  pair
  identification  domain  for  transfer  learning \cite{sun2019fine},  it's still
  quite  hard  to  perform  well  on  task-specific dataset because the training
  objective  of  sentence-pair  model  is semantic similarity, which is slightly
  different from the target of classification task. That is to say, there always
  exists  some intent classes which cannot be distinguished by each other merely
  depending on semantic similarity. For example, we have 2 user query saying, A:
  What  should I do if I want to change my password? B: What is the modification
  rule if I want to change my password? In this case, sentence A is semantically
  similar  to  B,  since  they  both  express  the desire for changing password.
  However, it's still reasonable to classify them into 2 intent classes, since A
  is asking for the method for changing password, while B is asking for the rule
  to follow when creating a new password.

  The  above  limitation  motivates  us  to  propose a joint system of both text
  classification  task  and  sentence-pair  semantic  similarity task, named SFC
  here, shown in Fig. \ref{fig:ver.1.3}. Our goal is to utilize the advantage of
  the  feasibility  of  transfer learning based on semantic similarity, while in
  the  mean time, add the classification task's target into the training process
  of            sentence-pair           model           for           multi-task
  learning \cite{caruana1993multitask,collobert2008unified}.  To  obtain  such  a
  joint  system,  we  first start with preparation work, which is pre-training a
  sentence-pair   model   on  external  corpus  for  sentence  pair  duplication
  identification,  and  then  find  an auxiliary model or tool which can help us
  sample out top-K most related intent classes. This auxiliary model or tool can
  either  be searching engine such as elasticsearch \cite{divya2013elasticsearch}
  or  a  text  classification  model  trained  on  task-specific  chatbot  data.
  Afterwards,  we  further  fine-tune  the  sentence-pair  model with 2 training
  tasks:  1)  the  regular  sentence pair similarity tasks. Here we use the text
  classification  model  as  an auxiliary model in this paper to sample sentence
  pairs       for       training       based      on      negative      sampling
  strategy \cite{bamler2020extreme}.  The  reason is that a task-specific chatbot
  often  has hundreds of intent classes, which forms too many sentence pairs for
  training  since  any  two  sentences  from  two  different  classes can form a
  sentence-pair     of     negative     sample.     Besides,     according    to
  Bamler \cite{bamler2020extreme},   training   on  negative  samples  from  most
  confusing  wrong  label  can  help  model  converge  faster  and obtain better
  performance.  2) the classification task on top-K (here K is a hyperparameter)
  candidate  classes  provided  by  our auxiliary model or tool. Here we use the
  average  pooling  of  the  representation  given  by  sentence-pair  model for
  sentence pairs formed by the user input query and the corpus sentences in each
  class  as  the  feature.  We  stack the task-specific layers on the top of the
  shared-parameter  sentence-pair  model structure, which is a classic parameter
  sharing                                   mechanism \cite{collobert2008unified,
  subramanian2018learning,liu2019multi}   for  multi-task \cite{sun2019fine}.  In
  this  way, specific knowledge contained in these 2 tasks can be fully explored
  and deeply interact with each other to obtain better overall performance.

@@

  3)  Experiment  results  on  4  public and 1 private short-text classification
  datasets with respect to intent recognition tasks under conversational chatbot
  scenario  all demonstrated that our proposed SFC joint system with multi-task,
  as  well  as  the  improved  version  of  SFC,  can  both  achieve  remarkable
  improvement comparing to some powerful single-task and single-model baselines,
  especially under few-shot settings.

@@

      \begin{tabular}{|c|c|c|c|c|c|}
        \hline 
        Dataset & Domain & \#Class & \#Samples & \#Samples/Class & Experimental Settings\tabularnewline
        \hline 
        ITG & FAQ chatbot & 228 & 3938 & 17 & 3-fold\tabularnewline
        \hline 
        Amazon-670k & Product review & 250 & 1942 & 8 & class sampling\tabularnewline
        \hline 
        HWU64 & Intent detection & 64 & [320, 640, 960, 1280] & [5, 10, 15, 20] & data sampling\tabularnewline
        \hline 
        CLINC150 & Intent detection & 150 & [750, 1500, 2250] & [5, 10, 15] & data sampling\tabularnewline
        \hline 
        BANKING77 & Intent detection & 77 & [385, 770, 1155, 1540] & [5, 10, 15, 20] & data sampling\tabularnewline
        \hline 
      \end{tabular}
      \par
    \end{centering}
    \caption{Key statistics for experimental datasets and few shot settings}

@@

  $\bullet$ \textbf{Subset of recently-released intent detection datasets}.

  From our point of view, task-specific chatbot usually starts with small amount
  of  data at the initial stage. As more and more customers start to communicate
  with  the  chatbot,  more data can be extracted from conversation log. Base on
  these  assumptions,  we  do  experiments  on  the following 3 intent detection
  dataset    under    sampling    method    similar    to    the   approach   in
   \cite{casanueva2020efficient}  but  in  more  sophisticated few-shot settings,
  while  keeping the same test sets for each experiment to see the change of the
  improvement scale with decreasing data amount.

@@

  \textbf{HWU64}   is  a  intent  detection  dataset  designed  for  home  robot
  scenario \cite{liu2019benchmarking}.  It aims at the specific task of capturing
  the  intent  for  different  user  requests  to  home  robot  and  finding the
  corresponding answer. The raw dataset contains 25,716 data points for 64 class
  labels  through  crowdsourcing,  which  raises  much  more  data  than  common
  real-world  circumstances. Therefore, we experiment with the setups where only
  5,  10,  15  and 20 training data points are available for each class label by
  random sampling.

  \textbf{CLINC150}  is a dataset designed for task-oriented systems with 23,700
  queries  that  are  short  and unstructured for 150 intents, in the same style
  made  by real users through crowdsourcing \cite{larson2019evaluation}. Here, we
  sample 5, 10 and 15 training data points for each intent class label by random
  sampling for our experimental setup.

  \textbf{BANKING77}  is  a intent detection dataset for bank customer services.
  The    raw    dataset    contains    13,084   data   points   for   77   class
  labels \cite{casanueva2020efficient}.  Here,  we  sample 5, 10, 15, 20 training
  data  points  for  each  class  label  by random sampling for our experimental
  setup.

@@

  \textbf{Group  1:  single-task  classification  model.} This group of baseline
  models  are  based  on the network structure for classification model shown in
  \ref{Fig.1.sub.1}.   As   for   non-pretrained   neural   network,  we  choose
  TextCNN \cite{kim2014convolutional}  as  a  representative baseline. As for the
  pretrained  model,  we choose RoBERTa, BERT and ALBERT, all of which are among
  the  most  powerful pretrained models that achieve outstanding results in many
  NLP  tasks,  in  our  experimental  setting. Since we apply RoBERTa in our SFC
  system,   we   naturally  adopt  single-task  classification  model  based  on
  RoBERTa-large  and RoBERTa-base as our baseline models. Besides, we also adopt
  the       recently       released       baseline       pre-trained       model
  ALBERT-xxlarge \cite{lan2019albert}  for  comparison.  Since  we  use different
  few-shot   settings   from    \cite{casanueva2020efficient}  for  those  intent
  detection  datasets,  we also use BERT-large tuned on our few-shot settings as
  an essential reference for experimental results.

@@

  \textbf{Analysis  under  few-shot  scenarios.}  Next,  we  analyze the overall
  improvement trend of SFC under various few-shot settings. Our main focus is to
  study chatbot building in common situation during real-world application where
  only  a  few  sample  sentences are available for each class. The experimental
  results  for  the  3  diverse  intent detection datasets (ClINC150, BANKING77,
  HWU64) under few-shot settings indicate that our proposed joint SFC can
  achieve  2.47  percents  average  improvement  over  one  of the most powerful
  baseline  models,  which  is  RoBERTa-large classification model, on all these
  few-shot  settings. Moreover, the improvement in F-score becomes more and more
  prominent  with the decrease of data sample number per class. Our joint
  SFC   can  achieve  4.97  percents  average  improvements  over  RoBERTa-large
  classification model in the most extreme few-shot setting with only 5 training
  samples  per class. This observation indicates that SFC is especially suitable
  for  being  applied during the initial stage of building task-specific chatbot
  system,  where  the  amount  of  data sample for each class label is extremely
  scarce.

  \textbf{Comparison  between  label  embedding  and  sentence  pair embedding.}
  Afterwards,  we  start  to  compare the performance of SFC with the text-label
  joint  embedding  baseline  named as LEAM. As shown in Table \ref{tbe:table2},
  joint  SFC outperforms LEAM by 14.25 percents on average F-score over 5
  datasets.  Besides,  although the labels within the 3 diverse intent detection
  datasets  (CLINC150,  BANKING77,  HWU64)  are  all  being  well-defined as the
  semantic  summary  of  each intent, joint SFC still outperforms LEAM by
  14.69  percents  on  average  F-score.  This  observation  indicates that it's
  usually  quite  hard  to  use  a  fixed  label  with  only  a few words as the
  representative  of  all  the  sentences  within a class since there are always
  plenty of ways for people to express questions or queries about the same topic
  during   a   conversation  with  chatbot.  Therefore,  learning  the  semantic
  similarity  between  user  query  and  sample  sentences through sentence pair
  embedding  can be more reliable than making decision according to the semantic
  relationship between user query and class labels through label embedding.

>>>>

  Task-specific  conversational  chatbots  \cite{wen2016network}  can be applied
  into  many practical products. For example, it can share the working burden of
  human  customer  service  agents  responsible  for  customers' questions about
  certain  products  or services. Another example is the speech-based control in
  devices  like  intelligent  speakers, such as Siri, Google home, Baidu Xiaodu,
  and smart home equipments. Regardless of whether these conversations belong to
  single-turn or multi-turn, an essential technique underlying the chatbot is to
  identify  the  real  intention  behind  a  user's  question or reply. Then the
  chatbot  is caple of conducting expected action, text reply or accomplish some
  operations.

  Although  task-specific  conversational chatbot already has a wide application
  in various business and industries, it's still a quite challenging task due to
  its  natural  properties of low-resource. First, customers' utterance within a
  conversation  is  usually  quite  short  and  completed  in a few words. Short
  text \cite{song2014short}  are generally more ambiguous in comparison with long
  texts  such  as  paragraphs  or  documents  since  they  don't  contain enough
  contextual  information,  which poses a great challenge \cite{chen2019deep} for
  classification        task \cite{phan2008learning,yan2009dynamic,hua2015short}.
  Second,  at  the  initial  stage  of  building  a chatbot for specific task or
  service, it's often extremely hard to collect sufficient data samples for each
  class.  The  reason  is  that it's usually too expensive and time consuming to
  extract  different  ways  of natural language expression for each intention class
  from  history  conversation  log  or  even  manually  compose  the corpus from
  nothing.  Therefore,  the  key  to  building a task-specific chatbot with high
  performance     becomes     solving     the     challenge     of    short-text
  classification \cite{sriram2010short}        problem       under       few-shot
  setting \cite{yu2018diverse}.

@@

  Despite the success of these 3 approaches, they still have some limitations in
  task-specific chatbot scenario. As for the text classification model approach,
  it's   quite   hard  to  use  data  augmentation  method  to  solve  the  data
  insufficiency  challenge since the it's usually unfeasible to get large amount
  of  domain-specific  data  when  facing  a  new  task. As for label-word joint
  embedding  approach,  it  poses  noticeable requirements on label information.
  However,  in  our  cases,  since  we have large amount of class labels without
  being well-defined for question answering chatbot, the only thing we can do is
  to  choose  one  of the user utterance as the class label to represent the key
  information  of  the  whole class. In this way, the label will become not only
  prolix  but  also  biased,  which  will  bring  some  negative effect to label
  embedding.  With  respect  to  the sentence-pair model approach, though we can
  obtain   large   amount  of  data  in  semantically  duplicate  sentence  pair
  identification  domain  for  transfer  learning \cite{sun2019fine},  it's still
  quite  hard  to  perform  well  on  task-specific dataset because the training
  objective  of  sentence-pair  model  is semantic similarity, which is slightly
  different from the target of classification task. That is to say, there always
  exists  some intention classes which cannot be distinguished by each other merely
  depending on semantic similarity. For example, we have 2 user query saying, A:
  What  should I do if I want to change my password? B: What is the modification
  rule if I want to change my password? In this case, sentence A is semantically
  similar  to  B,  since  they  both  express  the desire for changing password.
  However, it's still reasonable to classify them into 2 intention classes, since A
  is asking for the method for changing password, while B is asking for the rule
  to follow when creating a new password.

  The  above  limitation  motivates  us  to  propose a joint system of both text
  classification  task  and  sentence-pair  semantic  similarity task, named SFC
  here, shown in Fig. \ref{fig:ver.1.3}. Our goal is to utilize the advantage of
  the  feasibility  of  transfer learning based on semantic similarity, while in
  the  mean time, add the classification task's target into the training process
  of            sentence-pair           model           for           multi-task
  learning \cite{caruana1993multitask,collobert2008unified}.  To  obtain  such  a
  joint  system,  we  first start with preparation work, which is pre-training a
  sentence-pair   model   on  external  corpus  for  sentence  pair  duplication
  identification,  and  then  find  an auxiliary model or tool which can help us
  sample out top-K most related intention classes. This auxiliary model or tool can
  either  be searching engine such as elasticsearch \cite{divya2013elasticsearch}
  or  a  text  classification  model  trained  on  task-specific  chatbot  data.
  Afterwards,  we  further  fine-tune  the  sentence-pair  model with 2 training
  tasks:  1)  the  regular  sentence pair similarity tasks. Here we use the text
  classification  model  as  an auxiliary model in this paper to sample sentence
  pairs       for       training       based      on      negative      sampling
  strategy \cite{bamler2020extreme}.  The  reason is that a task-specific chatbot
  often  has hundreds of intention classes, which forms too many sentence pairs for
  training  since  any  two  sentences  from  two  different  classes can form a
  sentence-pair     of     negative     sample.     Besides,     according    to
  Bamler \cite{bamler2020extreme},   training   on  negative  samples  from  most
  confusing  wrong  label  can  help  model  converge  faster  and obtain better
  performance.  2) the classification task on top-K (here K is a hyperparameter)
  candidate  classes  provided  by  our auxiliary model or tool. Here we use the
  average  pooling  of  the  representation  given  by  sentence-pair  model for
  sentence pairs formed by the user input query and the corpus sentences in each
  class  as  the  feature.  We  stack the task-specific layers on the top of the
  shared-parameter  sentence-pair  model structure, which is a classic parameter
  sharing                                   mechanism \cite{collobert2008unified,
  subramanian2018learning,liu2019multi}   for  multi-task \cite{sun2019fine}.  In
  this  way, specific knowledge contained in these 2 tasks can be fully explored
  and deeply interact with each other to obtain better overall performance.

@@

  3)  Experiment  results  on  4  public and 1 private short-text classification
  datasets with respect to intention recognition tasks under conversational chatbot
  scenario  all demonstrated that our proposed SFC joint system with multi-task,
  as  well  as  the  improved  version  of  SFC,  can  both  achieve  remarkable
  improvement comparing to some powerful single-task and single-model baselines,
  especially under few-shot settings.

@@

      \begin{tabular}{|c|c|c|c|c|c|}
        \hline 
        Dataset & Domain & \#Class & \#Samples & \#Samples/Class & Experimental Settings\tabularnewline
        \hline 
        ITG & FAQ chatbot & 228 & 3938 & 17 & 3-fold\tabularnewline
        \hline 
        Amazon-670k & Product review & 250 & 1942 & 8 & class sampling\tabularnewline
        \hline 
        HWU64 & intention detection & 64 & [320, 640, 960, 1280] & [5, 10, 15, 20] & data sampling\tabularnewline
        \hline 
        CLINC150 & intention detection & 150 & [750, 1500, 2250] & [5, 10, 15] & data sampling\tabularnewline
        \hline 
        BANKING77 & intention detection & 77 & [385, 770, 1155, 1540] & [5, 10, 15, 20] & data sampling\tabularnewline
        \hline 
      \end{tabular}
      \par
    \end{centering}
    \caption{Key statistics for experimental datasets and few shot settings}

@@

  $\bullet$ \textbf{Subset of recently-released intention detection datasets}.

  From our point of view, task-specific chatbot usually starts with small amount
  of  data at the initial stage. As more and more customers start to communicate
  with  the  chatbot,  more data can be extracted from conversation log. Base on
  these  assumptions,  we  do  experiments  on  the following 3 intention detection
  dataset    under    sampling    method    similar    to    the   approach   in
   \cite{casanueva2020efficient}  but  in  more  sophisticated few-shot settings,
  while  keeping the same test sets for each experiment to see the change of the
  improvement scale with decreasing data amount.

@@

  \textbf{HWU64}   is  a  intention  detection  dataset  designed  for  home  robot
  scenario \cite{liu2019benchmarking}.  It aims at the specific task of capturing
  the  intention  for  different  user  requests  to  home  robot  and  finding the
  corresponding answer. The raw dataset contains 25,716 data points for 64 class
  labels  through  crowdsourcing,  which  raises  much  more  data  than  common
  real-world  circumstances. Therefore, we experiment with the setups where only
  5,  10,  15  and 20 training data points are available for each class label by
  random sampling.

  \textbf{CLINC150}  is a dataset designed for task-oriented systems with 23,700
  queries  that  are  short  and unstructured for 150 intents, in the same style
  made  by real users through crowdsourcing \cite{larson2019evaluation}. Here, we
  sample 5, 10 and 15 training data points for each intention class label by random
  sampling for our experimental setup.

  \textbf{BANKING77}  is  a intention detection dataset for bank customer services.
  The    raw    dataset    contains    13,084   data   points   for   77   class
  labels \cite{casanueva2020efficient}.  Here,  we  sample 5, 10, 15, 20 training
  data  points  for  each  class  label  by random sampling for our experimental
  setup.

@@

  \textbf{Group  1:  single-task  classification  model.} This group of baseline
  models  are  based  on the network structure for classification model shown in
  \ref{Fig.1.sub.1}.   As   for   non-pretrained   neural   network,  we  choose
  TextCNN \cite{kim2014convolutional}  as  a  representative baseline. As for the
  pretrained  model,  we choose RoBERTa, BERT and ALBERT, all of which are among
  the  most  powerful pretrained models that achieve outstanding results in many
  NLP  tasks,  in  our  experimental  setting. Since we apply RoBERTa in our SFC
  system,   we   naturally  adopt  single-task  classification  model  based  on
  RoBERTa-large  and RoBERTa-base as our baseline models. Besides, we also adopt
  the       recently       released       baseline       pre-trained       model
  ALBERT-xxlarge \cite{lan2019albert}  for  comparison.  Since  we  use different
  few-shot   settings   from    \cite{casanueva2020efficient}  for  those
  intention
  detection  datasets,  we also use BERT-large tuned on our few-shot settings as
  an essential reference for experimental results.

@@

  \textbf{Analysis  under  few-shot  scenarios.}  Next,  we  analyze the overall
  improvement trend of SFC under various few-shot settings. Our main focus is to
  study chatbot building in common situation during real-world application where
  only  a  few  sample  sentences are available for each class. The experimental
  results  for  the  3  diverse  intention detection datasets (ClINC150, BANKING77,
  HWU64) under few-shot settings indicate that our proposed joint SFC can
  achieve  2.47  percents  average  improvement  over  one  of the most powerful
  baseline  models,  which  is  RoBERTa-large classification model, on all these
  few-shot  settings. Moreover, the improvement in F-score becomes more and more
  prominent  with the decrease of data sample number per class. Our joint
  SFC   can  achieve  4.97  percents  average  improvements  over  RoBERTa-large
  classification model in the most extreme few-shot setting with only 5 training
  samples  per class. This observation indicates that SFC is especially suitable
  for  being  applied during the initial stage of building task-specific chatbot
  system,  where  the  amount  of  data sample for each class label is extremely
  scarce.

  \textbf{Comparison  between  label  embedding  and  sentence  pair embedding.}
  Afterwards,  we  start  to  compare the performance of SFC with the text-label
  joint  embedding  baseline  named as LEAM. As shown in Table \ref{tbe:table2},
  joint  SFC outperforms LEAM by 14.25 percents on average F-score over 5
  datasets.  Besides,  although the labels within the 3 diverse intention detection
  datasets  (CLINC150,  BANKING77,  HWU64)  are  all  being  well-defined as the
  semantic  summary  of  each intention, joint SFC still outperforms LEAM by
  14.69  percents  on  average  F-score.  This  observation  indicates that it's
  usually  quite  hard  to  use  a  fixed  label  with  only  a few words as the
  representative  of  all  the  sentences  within a class since there are always
  plenty of ways for people to express questions or queries about the same topic
  during   a   conversation  with  chatbot.  Therefore,  learning  the  semantic
  similarity  between  user  query  and  sample  sentences through sentence pair
  embedding  can be more reliable than making decision according to the semantic
  relationship between user query and class labels through label embedding.
